### 2017-09-20 profiles.Rmd
move pictures around and attempt to add rodent picture column to table

---
title: "Our Rodent Species"
---
<br>

```{r echo=FALSE, message=FALSE, warning=FALSE}

text_tbl <- data.frame(
  Rodents = c("![]('rodent_pictures/color_pictures/BA.png')",
              "![]('rodent_pictures/color_pictures/PB.png')",
              "![]('rodent_pictures/color_pictures/PH.png')",
              "![]('rodent_pictures/color_pictures/PI.png')",
              "![]('rodent_pictures/color_pictures/PP.png')",
              "![]('rodent_pictures/color_pictures/DM.png')",
              "![]('rodent_pictures/color_pictures/DO.png')",
              "![]('rodent_pictures/color_pictures/DS.png')",
              "![]('rodent_pictures/Rodent_silhouettes/5-White-throatedwoodrat.png')",
              "![]('rodent_pictures/Rodent_silhouettes/6-NorthernGrasshoppermouse.png')",
              "![]('rodent_pictures/Rodent_silhouettes/7-SouthernGrasshoppermouse.png')",
              "![]('rodent_pictures/color_pictures/PF.png')",
              "![]('rodent_pictures/color_pictures/PE.png')",
              "![]('rodent_pictures/color_pictures/PL.png')",
              "![]('rodent_pictures/color_pictures/PM.png')",
              "![]('rodent_pictures/Rodents_silhouettes/16-Fulvouseharvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/17-WesternHarvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/18-Plainsharvestmouse.png')",
              "![]('rodent_pictures/Rodents_silhouettes/19-Tawny-belliedcottonrat.png')",
              "![]('rodent_pictures/Rodents_silhouettes/20-HispidCottonRat.png')",
              "![]('rodent_pictures/Rodents_silhouettes/21-Yellownosedcottonrat.png)"), 
  Scientific.Name = c("_Baiomys taylori_", 
                      "_Chaetodipus baileyi_", 
                      "_Chaetodipus hispidus_",
                      "_Chaetodipus intermedius_",
                      "_Chaetodipus penicillatus_",
                      "_Dipodomys merriami_", 
                      "_Dipodomys ordii_", 
                      "_Dipodomys spectabilis_", 
                      "_Neotoma albigula_", 
                      "_Onychomys leucogaster_", 
                      "_Onychomys torridus_",
                      "_Perognathus flavus_",
                      "_Peromyscus eremicus_",
                      "_Peromyscus leucopus_",
                      "_Peromyscus maniculatus_",
                      "_Reithrodontomys fulvescens_",
                      "_Reithrodontomys megalotis_",
                      "_Reithrodontomys montanus_",
                      "_Sigmodon fulviventer_",
                      "_Sigmodon hispidus_",
                      "_Sigmodon ochorognathus_"),
  Common.Name = c("northern pygmy mouse", 
                  "Bailey's pocket mouse",
                  "hispid pocket mouse",
                  "rock pocket mouse",
                  "desert pocket mouse",
                  "Merriam's kangaroo rat", 
                  "Ord's kangaroo rat", 
                  "banner-tailed kangaroo rat", 
                  "white-throated woodrat", 
                  "northern grasshopper/scorpion mouse", 
                  "southern grasshopper/scorpion mouse",
                  "silky pocket mouse",
                  "cactus mouse",
                  "white-footed mouse",
                  "deer mouse",
                  "fulvous harvest mouse",
                  "western harvest mouse",
                  "plains harvest mouse",
                  "tawny-bellied cotton rat",
                  "white-bellied cotton rat",
                  "yellow-nosed cotton rat"),
  Description = c(
  "Yes, apparently southeastern Arizona counts as 'northern' for this genus. One of two competitors for title of 'smallest Portal rodent,' _B. taylori_ is currently an infrequent but welcome visitor to the site. As it is primarily a grassland species, individuals caught on the site rarely stay for long. Interestingly, the roughly 8-gram _B. taylori_ is often found in association with cotton rats (_Sigmodon_), which can weight in excess of 100 grams.", 
  "Bailey's pocket mouse made its first appearance on the site in the mid-1990s. It is essentially a giant version of its congeneric relative, the desert pocket mouse (_C. penicillatus_). While close in size to Merriam's and Ord's kangaroo rats, it is an inferior competitor and, therefore, is found almost solely on kangaroo rat exclosure plots.",
  "The hispid pocket mouse looks like a cactus mouse in pocket mouse form; it's the size of _C. baileyi_ with a beautiful lateral orange stripe. _C. hispidus_ has historically been the least frequently recorded species at the site. In the past few months, however, we've caught one or two each trapping period. Nevertheless, it remains a very transient speices; each individual caught this year has been caught once and then never seen again.",
  "_C. intermedius_ is another pocket mouse, though very rarely found at the site as it prefers rockier habitats; it is very similar to _C. penicillatus_ except in habitat preference.",
  "Although _C. penicillatus_ had been found at the site since the beginning of the study, it didn't become such an important player in the system until the 2000s. Now, this small pocket mouse is one of the main seasonal drivers of total rodent abundance at the site. The population explodes in the summers with known individuals, new adults, and brand new babies alike; they are ubiquitous across the site, regardless of plot type. In the winters, _C. penicillatus_ enters a type of torpor (not quite hibernation), creating a large seasonal fluctuation in abundance.  ",
  "Kangaroo rats, with their giant hind feet and namesake bipedal movement, are some of the most beloved species at the site. Of the three species in the _Dipodomys_ genus, _D. merriami_ is currently the most abundant and, presumably, dominant in the system.",
  "Ord's kangaroo rat is nearly indistinguishable from Merriam's kangaroo rat except for a tiny fifth toe on its hind feet that the Merriams lack. _D. ordii_ can be thought of as _D. merriami's_ kid brother who's always aspiring to be like his brother but can't quite keep up; a grass-loving species, _D. ordii_ is often just trailing _D. merriami_ in abundance.",
  "The largest of the three kangaroo rats at the site, _D. spectabilis_ (fondly referred to as 'spectabs' by those in the know) is known for its striking tail. It has by far the largest feet on the site at a whopping 48 mm average and often weighs in at over 100 grams, nearly twice the size of the other kangaroo rats. Spectabs were once running the show at Portal. As the site became shrubbier, however, the reign of _D. spectabilis_ came to a slow end in the 1990's. Since then, a few individuals have popped up here and there but haven't stuck around, often heading for grassier pastures.",
  "Along with being by far the largest rodent species at the site (my, what big teeth you have...), _Neotoma albigula_ also has a big reputation---roughly 200 grams worth. Woodrats are quite different ecologically from the rest of the species at the site. Unlike the majority of our rodents which are granivorous, _N. albigula_ is primarily a foliovore. (The resident woodrat in our ramada, however, seems to eat just about anything.) Woodrats---also commonly known as packrats---build middens, or large nests, out of sticks and any 'interesting' debris they can get their paws on. While their numbers at the site aren't prolific, research assistants know exactly which kangaroo rat exclosure plots this species tends to inhabit and secretly wish they'd brought leather gloves with them.",
  "Amongst smammal enthusiasts such as ourselves, the carnivorous grasshopper mice (_Onychomys_ genus) are legendary. Their name(s) and [this video](https://www.youtube.com/watch?v=ohd_mSIWTXk) basically say it all. In addition to eating insects and scorpions, they have been known to eat venomous centipedes and even other mice. Both species also have a distinctive high-pitched 'howl' when defending their territory. Mighty though they are, both species, and _O. leucogaster_ in particular, have hilariously short tails. For unknown reasons, the northern grasshopper mouse is less common at the site than the southern.",
  "Aside from overall range, the southern species differs little from the northern (_O. leucogaster_) except for a slightly longer tail and somewhat more elongated appearance. This species is more abundant at the site and can be found in all plot types.",
  "_Perognathus flavus_ rivals only _Baiomys taylorii_ for the title of smallest rodent at our site, weighing between 5 and 10 grams. Tiny, beautiful, and immensely soft, this aptly named silky pocket mouse is a favorite of previous and current research assistants alike. Its relatively low abundance and recent rarity makes it an especially exciting catch.",
  "Often having a bold lateral orange stripe, the cactus mouse is one of the most colorful and striking mice at Portal. It is also the most commonly found _Peromyscus_ species at the site, usually found in kangaroo rat exclosures and incidentally in the full removal plots. Unlike the white-footed mouse and deer mouse, however, it is mainly restricted to the deserts in the American Southwest and northern Mexico.",
  "This _Peromyscus_ species, while very common throughout the majority of the United States, is not all that common at Portal. _Peromyscus_ are good climbers and, therefore, frequently end up in the full rodent exclosure plots.",
  "Like _P. leucopus_, _P. maniculatus_ is nearly ubiquitous throughout the continental United States but is not a particularly common resident at the site. The desert variants of these two species are challenging to distinguish between, adding a fun challenge to trapping.",
  "The fulvous harvest mouse is an infrequent visitor to the site. It is the largest---and the rarest---harvest mouse species found at the site. Like many of the _Reithrodontomys_ caught at the site, it is most often found in kangaroo rat exclosures and full rodent removals.",
  "_R. megalotis_ is the most common harvest mouse found at Portal. Harvest mice abundance tends to remain fairly low throughout the site, though there is sometimes an increase in the winter when the desert pocket mice (_C. penicillatus_) are in torpor.",
  "While very similar to _R. megalotis_, _R. montanus_ is notably smaller than the other harvest mice at the site. It is less commonly found at the site than _R. megalotis_ but plays a similar role ecologically.",
  "A favorite of many Portal research assistants, rodents in the _Sigmodon_ genus are best described as 'squishy like cotton.' Okay, so maybe that's not the most descriptive identifier, but their size and coloration make them an easily identifiable genus. _S. fulviventer_ is one of the two more frequently seen cotton rat species at the site. It is distinguishable by the yellow (rather than white) fur on its belly, as indicated by its descriptive common name. ",
  "_S. hispidus_, with a white rather than a tawny belly, is the other cotton rat species that somewhat frequently has been caught at the site. The _Sigmodons_ seem to come through the site in waves; some years are boom years while others are bust years. What is driving these population cycles? How do _Sigmodons_ move through the regional landscape? We don't know but sure would like to find out.",
  "_S. ochorognathus_ almost seems like an urban (erm...rural) legend nowadays. A few individuals were caught in the 1990s, but none have been seen since."
  )
)

knitr::kable(text_tbl)

```