### 2019-02-27 evaluation.Rmd
Update evaluation.Rmd

---
title: "Evaluation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
  species <- rodent_spp(set = "evalplot")
  fcasts <- read_casts() %>%
            select_casts(species = species, level = "Controls") %>%
            append_observed_to_cast()
  fcast_errors <- measure_cast_error(fcasts)
  hcasts <- read_casts(cast_type = "hindcasts") %>%
            select_casts(species = species, level = "Controls") %>%
            append_observed_to_cast()
  hcast_errors <- measure_cast_error(hcasts)
```

## Metrics
**Error**: Error is the raw difference between a predicted point value ($\hat{y}_{i}$) and observed value ($y_{i}$): $e_{i} = \hat{y}_{i} - y_{i}$. Although we generally summarize errors in a fashion that results in non-negative values only (*e.g.*, via squaring), the raw error provides information about prediction bias (if the models tend to over- or under-predict in a systematic fashion).

**RMSE**: Root Mean Square Error (RMSE) is used to evaluate the absolute accuracy of the point estimates of a forecast. Across $N$ predicted values in a time series, $RMSE = \sqrt{\frac{\sum_{i=1}^{N}{e_{i}^2}}{N}}$.

**Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction. Because we use a 90% CI, coverage would ideally be equal to 0.90. If coverage is higher than 0.90, the forecast's intervals are too wide; if it's lower than 0.90, the forecast intervals are too narrow.

## How have the models done recently?

First, we evaluate the models' abilities to forecast the data since the beginning of the forecasting project.

## How would have the models done historically?

Here, we evaluate the models' abilities to hindcast the historic data.