### 2019-02-27 report.Rmd

---
title: "Report - Latest Rodent Data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
'%>%' <- magrittr::'%>%'
source("tools/forecast_tools.R")
```

Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
```{r, echo=FALSE}
plot_cast_point(with_census = TRUE)
```

```{r, echo=FALSE}
plot_cast_point(with_census = TRUE, level = "All")
```