### 2019-02-27 evaluation.Rmd
removing old top r chunks in rmds

---
title: "Evaluation"
---

## Metrics
**Error**: Error is the raw difference between a predicted point value ($\hat{y}_{i}$) and observed value ($y_{i}$): $e_{i} = \hat{y}_{i} - y_{i}$. Although we generally summarize errors in a fashion that results in non-negative values only (*e.g.*, via squaring), the raw error provides information about prediction bias (if the models tend to over- or under-predict in a systematic fashion).

**RMSE**: Root Mean Square Error (RMSE) is used to evaluate the absolute accuracy of the point estimates of a forecast. Across $N$ predicted values in a time series, $RMSE = \sqrt{\frac{\sum_{i=1}^{N}{e_{i}^2}}{N}}$.

**Coverage**: Coverage is used to evaluate the uncertainty in a forecast and is the percentage of observations which fell within the confidence interval of the prediction. Because we use a 90% CI, coverage would ideally be equal to 0.90. If coverage is higher than 0.90, the forecast's intervals are too wide; if it's lower than 0.90, the forecast intervals are too narrow.

## How have the models done recently?

This figure shows the raw error as a function of lead time (how far in advance the prediction was made) for three forecast dates (colors of lines) across each combination of model (column) and species (row) for a selection of historically common species and the total rodents, for the control (unmanipulated) plots only.

```{r recent_eval, echo=FALSE, fig.width=10, fig.height=10}
plot_err_lead_spp_mods()
```

## How would have the models done historically?

Here, we evaluate the models' abilities to hindcast the historic data.