### 2023-04-26 app/models.Rmd
setup

---
title: "Models"
bibliography: refs.bibtex
output: 
  rmarkdown::html_document 
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(collapse = TRUE,
                        comment  = "#>")

  models_controls <- prefab_models_controls( )
  models_names    <- unlist(lapply(lapply(models_controls, getElement, "metadata"), getElement, "print_name"))
  nmodels         <- length(models_names)
  nmodels_text    <- as.character(english(nmodels))

```

<br>

We currently analyze and forecast rodent data at Portal using `r nmodels_text` models: `r models_names` [@PortalPredictions].

```{r echo = FALSE, results = "asis"}

  template <- "## %s \n \n %s \n \n   \n"

  for (i in 1:nmodels) {
    model_i_controls <- models_controls[[i]]
    cat(sprintf(template, model_i_controls$metadata$print_name, model_i_controls$metadata$text))
  }
```

# References